WIDS 2024 Final Presentation

### 2024-01-12 Documents/Paper.qmd
updates in code presentation and paper

format: 
  pdf:
    number-sections: true
    toc: false
    geometry: 
      - top=3cm
date: 2024-01-12
date-format: iso
documentclass: article
editor: source
abstract: "This is our informative abstract of fewer than 200 words. It describes what we investigate, how we investigate it, and what we find."
#bibliography: main.bib
---

@@

Unions play a crucial role in negotiating wages and benefits for workers through collective bargaining agreements. The political landscape, particularly the House of Representatives, holds significant legislative authority in influencing workers' rights and labor policies. We examine the relationship between union employment, the composition of the House delegation, and economic variables such as income, minimum wages, and union wages.

In our research, we seek to understand how union membership and economic factors influence the level of representation by the Democratic Party in the House of Representatives. We hypothesize that there will be an association between the Democratic Party and the employment of union and collective bargaining employees.

>>>>

format: 
  pdf:
    number-sections: true
    toc: false
    geometry: 
      - top=3cm
date: 2024-01-12
date-format: iso
documentclass: article
editor: source
abstract: "For our project, our goal is to examine the relationships between collective bargaining agreements, U.S. house election results, and incomes/wages of Americans. The motivation for this project is the rise in wealth inequality and populism in America and the recent surge in union activity. In our project, we find that all three metrics are related. Using regression testing we show that for unions and political leanings, this relationship has surged and deepened in the last four decades from 1984 to 2019"
#bibliography: main.bib
---

@@

Unions play a crucial role in negotiating wages and benefits for workers through collective bargaining agreements. The political landscape, particularly the House of Representatives, holds significant legislative authority in influencing workers' rights and labor policies. We examine the relationship between unions and collective bargaining agreements, U.S. House of Representatives election results, and economic variables including the incomes and wages of Americans.

In our research, we seek to understand how union membership and economic factors influence the level of representation by the Democratic Party in the House of Representatives. We hypothesize that there will be an association between the proportion of the House delegation that is Democratic and the employment of union and collective bargaining workers.

### 2024-01-12 Documents/Presentation.qmd
updates in code presentation and paper

For our project, our goal was to examine the relationships between collective bargaining agreements, U.S. house election results, and incomes/wages of Americans. The motivation for this project is the rise in wealth inequality and populism in America and the recent surge in union activity. In our project, we found that all three metrics are related. Using regression testing we show that for unions and political leanings, this relationship has surged and deepened in the last forty years.

@@

-   **Union data**
    -   Gathered from Tidy Tuesday, although the original data comes from the Union Membership and Coverage Database at Georgia state university.
    -   For our project, the relevant tables include one that details union membership counts by state by year and another that tracks union and non union wages by year.
-   **House election data**
    -   Also gathered from Tidy Tuesday's repository, this data originally came from the MIT Election Data and Science Lab.
    -   The variables we selected from the data set were year, state, district, and election winner

@@

-   All of our data had undergone basic cleaning before we downloaded it so all of the columns were correctly formatted and there was almost 0 NA values.
-   each data set was pared down so that only the variables relevant to our project was completed.
-   for our house election data, the table was manipulated so that only rows of candidates who won their election remained.
    -   other minor changes included renaming variables and recoding a few outlier observations in the `party` variable.
-   For the union data, we filtered the datasets down to not include union sectors (ex. construction-unions) and added columns to adjust for inflation.
-   For our wage data, we filtered out race, as this was outside of the scope of our study.

@@

-   Our last data manipulation was combining all of our tables into two data sets: one that stores data at a national level, and another that stores data at a state level.
-   some of the variables were only provided at the state level so they were summarized and added to the national data set.

@@

::: {.column width="40%"}
![Plot 6](./Graph_Images/CB_by_decade1.png){width=80%}
![Plot 7](./Graph_Images/CB_by_decade2.png){width=80%}

:::
:::

## Models {.smaller}
::: columns
::: {.column width="60%"}
-   Our regression analysis shows that the south and north central regions in the united states are the main regions that have seen union membership rates and political leanings show increased correlation.
- T-test between union and non union wages
  - 1980s: t = 22.846, df = 9.4096, p-value = 1.462e-09
    - mean of x = 26.306;  mean of y = 20.375
    - 95% CI: (5.347, 6.514)
  - 2010s: t = 22.846, df = 9.4096, p-value = 1.462e-09
    - mean of x = 3.786;  mean of y = 25.664
    - 95% CI: (3.786, 5.952)
:::

::: {.column width="40%"}
![Plot 8](./Graph_Images/Regression_plot.png)

:::
:::

@@

From our data analysis it can be infered that collective bargaining agreements have become a very political concept. The charts in our exploratory data analysis and regression models show this. The relationship between unions and wages/income in the United States is much harder to draw inferences. While our t-tests and some charts show that unions can help workers earn more, other graphs show that unions have a negative effect on incomes.

## Where we would take our project in the future {.smaller}
- Wage data at the state level
- Data past 2019
- A greater amount and complexity of models

>>>>

For our project, our goal is to examine the relationships between collective bargaining agreements, U.S. house election results, and incomes/wages of Americans. The motivation for this project is the rise in wealth inequality and populism in America and the recent surge in union activity. In our project, we find that all three metrics are related. Using regression testing we show that for unions and political leanings, this relationship has surged and deepened in the last forty years.

@@

-   **Union data**
    -   Gathered from Tidy Tuesday, although the original data comes from the Union Membership and Coverage Database at Georgia State University.
    -   For our project, the relevant tables include one that details union membership counts by state by year and another that tracks union and non union wages by year.
-   **House election data**
    -   Also gathered from Tidy Tuesday's repository, this data originally came from the MIT Election Data and Science Lab.
    -   The variables we selected from the data set were year, state, district, and election winner

@@

-   All of our data had undergone basic cleaning before we downloaded it so all of the columns were correctly formatted and there was almost 0 NA values.
-   Each data set was pared down so that only the variables relevant to our project was completed.
-   For our house election data, the table was manipulated so that only rows of candidates who won their election remained.
    -   Other minor changes included renaming variables and recoding a few outlier observations in the `party` variable.
-   For the union data, we filtered the datasets down to not include union sectors (ex. construction-unions) and added columns to adjust for inflation.
-   For our wage data, we filtered out race, as this was outside of the scope of our study.

@@

-   Our last data manipulation was combining all of our tables into two data sets: one that stores data at a national level, and another that stores data at a state level.
-   Some of the variables were only provided at the state level so they were summarized and added to the national data set.

@@

::: {.column width="40%"}
![Plot 6](./Graph_Images/CB_by_decade1.png){width="80%"} ![Plot 7](./Graph_Images/CB_by_decade2.png){width="80%"}
:::
:::

## Models {.smaller}

::: columns
::: {.column width="60%"}
-   Our regression analysis shows that the south and north central regions in the United States are the main regions that have seen union membership rates and political leanings show increased correlation.
-   T-test between union and non union wages
    -   1980s: t = 22.846, df = 9.4096, p-value = 1.462e-09
        -   mean of x = 26.306; mean of y = 20.375
        -   95% CI: (5.347, 6.514)
    -   2010s: t = 22.846, df = 9.4096, p-value = 1.462e-09
        -   mean of x = 3.786; mean of y = 25.664
        -   95% CI: (3.786, 5.952)
:::

::: {.column width="40%"}
![Plot 8](./Graph_Images/Regression_plot.png)
:::
:::

@@

From our data analysis it can be inferred that collective bargaining agreements have become a very political concept. The charts in our exploratory data analysis and regression models show this. The relationship between unions and wages/income in the United States is much harder to draw inferences. While our t-tests and some charts show that unions can help workers earn more, other graphs show that unions have a negative effect on incomes.

## Where we would take our project in the future {.smaller}

-   Union wage data at the state level
-   Data past 2019
-   A greater amount and complexity of models
